---
title: "Untitled"
author: "Ethan Caldecott"
date: "2022-11-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(dplyr)
library(tidyselect)
library(tidyverse)
library(lubridate)
library(rvest)
library(stringr)
library(scales)
library(plotly)
library(rpart)
library(rpart.plot)
```


#Read in Contract Data
```{r}
CurrentContractData <- read_csv("2023contract.txt")
#via https://www.basketball-reference.com/contracts/players.html
```

#Clean Contract Data Set
```{r}
CurrentContractData$`2022-23` = as.numeric(gsub("[\\$,]", "", CurrentContractData$`2022-23`))
CurrentContractData$`2023-24` = as.numeric(gsub("[\\$,]", "", CurrentContractData$`2023-24`))
CurrentContractData$`2024-25` = as.numeric(gsub("[\\$,]", "", CurrentContractData$`2024-25`))
CurrentContractData$`2025-26` = as.numeric(gsub("[\\$,]", "", CurrentContractData$`2025-26`))
CurrentContractData$`2026-27` = as.numeric(gsub("[\\$,]", "", CurrentContractData$`2026-27`))
CurrentContractData$`2027-28` = as.numeric(gsub("[\\$,]", "", CurrentContractData$`2027-28`))
CurrentContractData$`Guaranteed` = as.numeric(gsub("[\\$,]", "", CurrentContractData$`Guaranteed`))

CurrentContractData <- CurrentContractData %>% distinct(Player, .keep_all = TRUE)

CurrentContractData %>% 
  select(Player, Tm, `2022-23`, `2023-24`, `2024-25`, `2025-26`, `2026-27`, `2027-28`,`Guaranteed`) %>% 
  arrange(`2022-23`)

GuaranteedContract <- CurrentContractData %>% select(Player, Tm, Guaranteed) %>% arrange((Guaranteed))
GuaranteedContract

# 469 without duplicates
```

# 2022-2023 Season Stats
```{r}
CurrentPlayerData <- read_csv("2023bball.txt")
#via https://www.basketball-reference.com/leagues/NBA_2023_totals.html


CurrentPlayerData 
#454 rows

```

# 2021-2022 Season Stats
```{r}
PastPlayerData <-read_csv("2022bball.txt")

#via https://www.basketball-reference.com/leagues/NBA_2022_totals.html

PastPlayerData 
#605 players
```

# Inner Join
```{r}
CurrentxContract <- CurrentPlayerData %>% inner_join(CurrentContractData, by = 'Player')
CurrentxContract
#414 players

ContractxCurrent <- CurrentContractData %>% inner_join(CurrentPlayerData, by = 'Player')
ContractxCurrent 
  
#414 players
```
# Explaining Missing Values
## Anonamaly ContractDataBadValue: 1
```{r}
# 1 two way with guranteed moneay in the contract dataset
CurrentContractData %>% 
  select(Player, Tm, `2022-23`) %>%
  filter(is.na(`2022-23`))
#1 player in contract data set has no contract data but shows up in contract data set (Ish Wainwright is now a 2 way player guranteed money left over from last season)
```

## Explain Missing Values: 40
```{r}
CurrentLContract <- CurrentPlayerData %>% left_join(CurrentContractData, by = 'Player') 
CurrentLContract #total of 454 players with stats

CurrentLContract[is.na(CurrentLContract$`2022-23`),] 
#40 of the players with NA contract values for their 2022-23 are on two-way contracts

#Non-two way players
CombineData <- CurrentLContract[!is.na(CurrentLContract$`2022-23`),]
CombineData #414 rostered players with stats and contracts
```

## Explain Missing Values cont  
```{r}
Cdf <- CurrentContractData %>% left_join(CurrentPlayerData, by = 'Player') %>% distinct(Player, .keep_all = TRUE)
Cdf
```


### Explain Missing Value: 55
```{r}
#55 player have 0 games played and do not show up on the stats for 2023
DNPCdf <- Cdf %>%
  filter(is.na(G)) %>% 
  select(c(2:10))
DNPCdf
```

#### Explaining Missing Values in the 55: 48
```{r}
# 46 of the 57 have stats from last season
Updated46 <- DNPCdf %>%
  left_join(PastPlayerData, by = 'Player') %>%
  filter(!is.na(Pos))

Updated46$Age <- Updated46$Age + 1 
#add 1 to age because it is last year

Updated46[ , colSums(is.na(Updated46)) < nrow(Updated46)] #eliminates all columns that are all NA


```

#### Explaining Missing Values in the 55: 9
```{r}
# 9  of the 57 who have no stats last season
Manual <- DNPCdf %>%
  left_join(PastPlayerData, by = 'Player') %>%
  filter(is.na(Pos)) 

#manually input data from basketball reference for each player
Manual$Age = c(25, 20, 32, 20, 29, 19, 34, 27, 28)
Manual$Pos = c("PF", "PF", "PF", "PG", "SF", "SF", "SF", "SF", "PG") 
Manual$Tm.y <- Manual$Tm.x
```

#### Explaining Missing Values in the 55: in the 9: DELETE 4
```{r}
#4 of the 9 are no longer in the NBA and were cut but had their contract restructured in a way where they are still getting payed
Manual1 <- Manual %>%
  filter(Player != "Andrew Nicholson" & Player != "Kyle Singler" & Player != "Troy Williams" & Player != "Demetrius Jackson") 

```

#### Explaining Missing Values in the 55: in the 9: Keep 5
```{r}
#5 of the 9 are rookies or vets that were injured all of last year and are yet to play
Manual1 
```

# Add the 46 and the 5 and add to the 414 
```{r}
P414 <- ContractxCurrent %>%
  select(c(2:10, 13:15))


E46 <- Updated46 %>% 
  select(1:9,11:13) %>%
  rename(Tm.y = Tm) 
E46
#5 didn't play last season or this
E5 <- Manual1 %>%
  select(1:9, 11:12,40) %>%
  rename()
E5

E51 <- E46 %>% full_join(E5)

F465 <- E51 %>% full_join(P414)
F465 #Official Data Set
```

# Clean F465 Data
## Delete the 1 and 1 new ones!
```{r}
F465
F465[F465 == 27431078] <- 9165471
F465

NBAData <- F465 %>%
  arrange(desc(`2022-23`)) %>%
  mutate(`2023-24` = replace_na(`2023-24`, 0))  %>%
  mutate(`2024-25` = replace_na(`2024-25`, 0))  %>%
  mutate(`2025-26` = replace_na(`2025-26`, 0))  %>%
  mutate(`2026-27` = replace_na(`2026-27`, 0))  %>%
  mutate(`2027-28` = replace_na(`2027-28`, 0))  %>%
  mutate(Guaranteed = replace_na(Guaranteed, 0))  %>%
  mutate(PercentGuaranteed = Guaranteed/(`2022-23` + `2023-24` + `2024-25` + `2025-26` +`2026-27`+ `2027-28`)*100) %>% 
  arrange(desc(PercentGuaranteed)) %>%
  select(1, 3:13) %>%
  mutate(Pos = str_replace_all(Pos, pattern = "SG-PG", replacement = "PG")) %>%
  rename(Team = Tm.y) %>%
  filter(1000000 < `2022-23` | `2023-24` != 0) %>%
  filter(`2022-23` != `2023-24` | `2022-23` > 1000000) %>%
  arrange(`2022-23`) %>%
  filter(Player != "Didi Louzada", Player != "Ish Wainright") #These players have dead cap in their salary or were cut
NBAData
```

View(NBAData)
# Analysis
Per Position Average of 2022-23 Salary of Rostered NBA Players
```{r}
NBAData %>% 
  group_by(Pos) %>%
  mutate(NumberOFPlayers = n()) %>%
  mutate(AverageSalary2223 = mean(`2022-23`)) %>%
  distinct(Pos, AverageSalary2223, NumberOFPlayers) %>%
  arrange(desc(AverageSalary2223))

# 11,209,354 is the highest position average salary for official rostered nba players.
```

Per Age Average of 2022-23 Salary of Rostered NBA Players
```{r}
NBAData %>% 
  group_by(Age) %>%
  mutate(NumberOFPlayers = n()) %>%
  mutate(AverageSalary2223 = mean(`2022-23`)) %>%
  distinct(Age, AverageSalary2223, NumberOFPlayers) %>%
  arrange(desc(AverageSalary2223))
# Age probably not a good metric as low number on average


```


Per Age & Position Average of 2022-23 Salary of Rostered NBA Players
```{r}
NBAData %>% 
  mutate(Age1 = cut(Age, breaks = c(18,20,22,24,26,28,30,32,34,36,43), labels = c("19-20", "21-22", "23-24", "25-26", "27-28", "29-30", "31-32", "33-34", "35-36", "37+"))) %>%
  #only 2 players that are above 38, 1 39 & 1 42
  select(Age, Age1) %>% 
  arrange(desc(Age)) %>%
  group_by(Age1) %>%
  summarize(n = n())
```



Average Salary of Rostered NBA Players
```{r}
mean(NBAData$`2022-23`)
```

Average Percent of Guaranteed Contracts of Rostered NBA Players
```{r}
mean(NBAData$PercentGuaranteed)
```
# Total $ of NBA Contracts
```{r}
sum(NBAData$`2022-23`)
```

# Number of Players
```{r}
count(NBAData)
```

```{r}
max(NBAData$`2022-23`)
min(NBAData$`2022-23`)
```

Per Age & Position Average of 2022-23 Salary of Rostered NBA Players
```{r, message = FALSE}
P <- NBAData %>%
  mutate(Age1 = cut(Age, breaks = c(18,20,22,24,26,28,30,32,34,43), labels = c("19-20", "21-22", "23-24", "25-26", "27-28", "29-30", "31-32", "33-34", "35+"))) %>%
  group_by(Age1, Pos) %>%
  summarize(AverageSalary2223 = mean(`2022-23`)) %>% 
  ggplot(aes(x = Age1, y= `AverageSalary2223`/1000000, color = Pos, group=Pos, text = paste("Average Salary:", round(AverageSalary2223/1000000, 1), "million USD", "\nPosition:", Pos, "\nPlayer's Age Range:", Age1, "years old"))) +
  geom_point() +
  geom_smooth(se = FALSE) + 
  scale_y_continuous(labels = comma_format(big.mark = ",")) + 
  labs(y = "Average Salary (million USD)", x = "Player's Age Range", title = "Average Player Salary by Position and Age", subtitle = "via NBA players in the 2022-2023 season with a Standard NBA Contract", color = "Player's Position", caption = "data from basketball-reference.com") +
  theme_classic() 
ggplotly(P, tooltip = c('text'))
P
```

# STATS

```{r}

XXX <- NBAData %>%
  rename(Pos2223 = Pos) %>%
  rename(Age2223 = Age) %>%
  inner_join(PastPlayerData, by = "Player")
CTest <- XXX %>%
  select(-Player, -Team, -Pos, -Age, -Tm, -`Player-additional`)

PGdata <- CTest %>% 
  filter(Pos2223 == "PG") %>%
  select(-Pos2223)

SGdata <- CTest %>% 
  filter(Pos2223 == "SG") %>%
  select(-Pos2223)

SFdata <- CTest %>% 
  filter(Pos2223 == "SF") %>%
  select(-Pos2223)

PFdata <- CTest %>% 
  filter(Pos2223 == "PF") %>%
  select(-Pos2223)

Cdata <- CTest %>% 
  filter(Pos2223 == "C") %>%
  select(-Pos2223)

Alldata <- CTest %>% select(-Pos2223)

```
 
```{r}
v = .5
corALL<- cor(Alldata) %>%
  as.data.frame() %>%
  mutate(var1 = rownames(.)) %>%
  gather(var2, value, -var1) %>%
  arrange(desc(value)) %>%
  filter(var1 == "2022-23") %>%
  filter(var2 != "2022-23" & var2 != "2023-24" & var2 != "Guaranteed" & var2 != "2024-25" & var2 != "2025-26" & var2 != "2026-27" & var2 != "2027-28")%>%
  filter(abs(value) >= v)
  
corPG <- cor(PGdata) %>%
  as.data.frame() %>%
  mutate(var1 = rownames(.)) %>%
  gather(var2, value, -var1) %>%
  arrange(desc(value)) %>%
  filter(var1 == "2022-23") %>%
  filter(var2 != "2022-23" & var2 != "2023-24" & var2 != "Guaranteed" & var2 != "2024-25" & var2 != "2025-26" & var2 != "2026-27" & var2 != "2027-28")%>%
  filter(abs(value) >= v)
  

corSG <- cor(SGdata) %>%
  as.data.frame() %>%
  mutate(var1 = rownames(.)) %>%
  gather(var2, value, -var1) %>%
  arrange(desc(value)) %>%
  filter(var1 == "2022-23") %>%
  filter(var2 != "2022-23" & var2 != "2023-24" & var2 != "Guaranteed" & var2 != "2024-25" & var2 != "2025-26" & var2 != "2026-27" & var2 != "2027-28")%>%
  filter(abs(value) >= v)
  

corSF <- cor(SFdata) %>%
  as.data.frame() %>%
  mutate(var1 = rownames(.)) %>%
  gather(var2, value, -var1) %>%
  arrange(desc(value)) %>%
  filter(var1 == "2022-23") %>%
  filter(var2 != "2022-23" & var2 != "2023-24" & var2 != "Guaranteed" & var2 != "2024-25" & var2 != "2025-26" & var2 != "2026-27" & var2 != "2027-28")%>%
  filter(abs(value) >= v)
  

corPF <- cor(PFdata) %>%
  as.data.frame() %>%
  mutate(var1 = rownames(.)) %>%
  gather(var2, value, -var1) %>%
  arrange(desc(value)) %>%
  filter(var1 == "2022-23") %>%
  filter(var2 != "2022-23" & var2 != "2023-24" & var2 != "Guaranteed" & var2 != "2024-25" & var2 != "2025-26" & var2 != "2026-27" & var2 != "2027-28")%>%
  filter(abs(value) >= v)
  

corC <- cor(Cdata) %>%
  as.data.frame() %>%
  mutate(var1 = rownames(.)) %>%
  gather(var2, value, -var1) %>%
  arrange(desc(value)) %>%
  filter(var1 == "2022-23") %>%
  filter(var2 != "2022-23" & var2 != "2023-24" & var2 != "Guaranteed" & var2 != "2024-25" & var2 != "2025-26" & var2 != "2026-27" & var2 != "2027-28") %>%
  filter(abs(value) >= v)
  
```

```{r}
PGdata %>%
  ggplot(aes(y = `2022-23`, x  = FT)) +
  geom_point() +
  geom_smooth(method = "lm")
```



```{r}
XYZ <- XXX %>%
  select(-Player, -`2023-24`,-`2024-25`, -`2025-26`, -`2026-27`, -`2027-28`, -Guaranteed, -PercentGuaranteed, -Rk, -`Player-additional`, -`Tm`, -Team, -Age) %>%
  filter(Pos == "PG")
XXX
XYZ

#HI1 <- rpart(`2022-23` ~ `2022-23` + `Pos2223`+ `Age2223`, `Pos`, `G`, `GS`, `MP`, `FG`, `FGA`, `FG%`, `3P`, `3PA`,`3P%`, `2P`, `2PA`, `2P%`, `eFG%`, `FT`, `FTA`, `FT%`, `ORB`, `DRB`, `TRB`, `AST`, `STL`, `BLK`, `TOV`, `PF`, `PTS`, data = XXX)

HI <- rpart(`2022-23` ~ ., data = XYZ)
rpart.plot(HI, type = 3, clip.right.labs = FALSE, branch = .3, under = TRUE)
rpart.rules(HI, cover = TRUE)
XYZ$predict <- predict(HI)
XYZ$resid <- round(resid(HI), digits = 0)

ggplot(XYZ,aes(y = resid, x = `2022-23`,color= resid > 0)) + geom_point()
```

```{r}
#Fair if 2022-23 salary is within 10% of the predict
OUF <- XYZ %>%
  filter(Pos == "PG") %>%
  mutate(Type = case_when(
    resid > .05 *predict ~ "Over", 
    resid < -.05*predict ~ "Under",
    TRUE ~ "Fair"
  )) %>%
  arrange(resid)
OUF
XXX

OUFnames <- OUF %>% 
  inner_join(XXX, by = c("2022-23", "Pos2223", "Age2223", "Pos", "G", "GS", "MP", "FG", "FGA", "FG%", "3P", "3PA", "3P%", "2P", "2PA", "2P%", "eFG%", "FT", "FTA", "FT%", "ORB", "DRB", "TRB", "AST", "STL", "BLK", "TOV", "PF", "PTS")) %>%
  select(`Player`, `2022-23`, `Team`, `Pos`, `Age2223`, `G`, `GS`, `MP`, `FG`, `FGA`, `FG%`, `3P`, `3PA`,`3P%`, `2P`, `2PA`, `2P%`, `eFG%`, `FT`, `FTA`, `FT%`, `ORB`, `DRB`, `TRB`, `AST`, `STL`, `BLK`, `TOV`, `PF`, `PTS`, `predict`, `resid`, `Type`)
  
OUFnames


```

# Graph Sorting

```{r}
NBArookies <- read_csv("C:/Users/ethan/OneDrive/Macalester Sophmore Fall/Introduction to Data Science/NBArookies.txt")
```
```{r}
fitModel <- function(df){
  mod <- df %>% 
      select(-Player, -`2023-24`,-`2024-25`, -`2025-26`, -`2026-27`, -`2027-28`, -Guaranteed, -PercentGuaranteed, -Rk, -`Player-additional`, -`Tm`, -Team, -Age) %>%
    rpart(`2022-23` ~ ., data = .) 
  
  mod
}


model_results <- XXX %>%
  group_by(Pos2223) %>%
  nest() %>%
  mutate(model = purrr::map(.x = data, ~fitModel(.x))) %>%
  mutate(predict = purrr::map(.x = model, ~predict(.x))) %>%
  mutate(resid = purrr::map(.x = model, ~resid(.x))) %>%
  select(-model) %>%
  unnest() %>%
  mutate(Type = case_when(
    resid > .2 *predict ~ "Overpaid", 
    resid < -.2*predict ~ "Underpaid",
    TRUE ~ "Fairly Paid"
  ))

model_results %>% 
  group_by(`Team`, `Type`) %>%
  summarize(NumberPlayers = n())

RegPlayer <- model_results %>%
  select("Player", "Pos2223", "Age2223", "Team","predict", "resid", "Type","2022-23", "2023-24", "2024-25","2025-26", "2026-27", "2027-28", "Guaranteed", "PercentGuaranteed", "G","GS","MP","FG","FGA", "FG%", "3P", "3PA", "3P%", "2P", "2PA", "2P%", "eFG%", "FT", "FTA", "FT%", "ORB", "DRB", "TRB", "AST", "STL", "BLK","TOV", "PF", "AST", "STL", "BLK","TOV","PF","PTS") %>%
  rename("Pos" = "Pos2223") %>%
  rename("Age" = "Age2223")

 
```

```{r}
#Players non-two way
#NBAData

#All players with 2021-2022 stats
#RegPlayer

#All player with no 2021-2022 stats that are not rookies
IRplayers <- NBAData %>% anti_join(model_results, by = "Player") %>% anti_join(NBArookies, by = "Player")
IRplayers$Type = "IR"
IRplayers <- IRplayers %>% 
   select("Player", "Pos", "Age", "Team","Type","2022-23", "2023-24", "2024-25","2025-26", "2026-27", "2027-28", "Guaranteed", "PercentGuaranteed")

#Rookies with no contract data (17 2nd round rookies signed to two-way contracts or cut)
CutRookies <- NBArookies %>%
  anti_join(NBAData, by = "Player") %>%
  select(Player, Tm)

#Confirm that we only have contract data for 41 of 58 rookies
#Rookies
Rookies <- NBAData %>% inner_join(NBArookies, by = "Player") %>%
  select(Player, Team, Pos)
Rookies$Type = "Rookies"

#Standard Contract Players + IR players

TOTdataset <- RegPlayer %>% 
  full_join(Rookies) %>%
  full_join(IRplayers) %>%
  filter(Team == "TOT")
TOTdataset$Team = c("GSW","LAL", "DAL", "IND")

NoTOTdataset <- RegPlayer %>% 
  full_join(Rookies) %>%
  full_join(IRplayers) %>%
  filter(Team != "TOT")


FinalDataSet <- TOTdataset %>% full_join(NoTOTdataset)
PosMakeUp <- FinalDataSet %>% 
  group_by(Team, Pos) %>%
  summarize(n = n()) %>%
  pivot_wider(names_from = Pos, values_from = n) 

uuuu <- PosMakeUp %>%
  left_join(updateabb_team) %>%
  mutate(Team = `Official Team Name`) %>%
  select(-`Official Team Name`)

```
```{r}
GraphData <- FinalDataSet %>%
  group_by(Team, Type) %>%
  summarize(Typetotal = n()) %>%
  group_by(Team) %>%
  mutate(Teamtotal = sum(Typetotal)) %>%
  arrange(Team)
  
GraphData %>% ggplot(aes(x=Team, y=Typetotal, fill=Type)) +
  geom_col() +
  coord_flip() +
  ylab("Number of Players")+
  labs(title = "Roster Construction of Each NBA Team in 2022-23")



```

```{r}
abb_team <- read_delim("C:/Users/ethan/OneDrive/Macalester Sophmore Fall/Introduction to Data Science/abb-team.txt", 
    delim = "-", escape_double = FALSE, col_names = FALSE,
    trim_ws = TRUE)

updateabb_team <- abb_team %>%
  rename("Team" = X1) %>%
  rename("Official Team Name" = X2)
updateabb_team

Combo1 <- GraphData %>%
  left_join(updateabb_team) %>%
  mutate(Team = `Official Team Name`) %>%
  select(-`Official Team Name`)
Combo1
```

```{r}
TeamRank <- read_csv("C:/Users/ethan/OneDrive/Macalester Sophmore Fall/Introduction to Data Science/TeamRank.txt") %>% rename(Win=W)
TeamRank
```

```{r}
Everything <- Combo1 %>% 
  left_join(TeamRank)

Save <- Everything %>% 
  pivot_wider(names_from = `Type`, values_from = `Typetotal`) %>%
  mutate("IR" = replace(IR, is.na(IR), 0)) %>%
  mutate("Fairly Paid" = replace(`Fairly Paid`, is.na(`Fairly Paid`), 0)) %>%
  mutate("Rookies" = replace(`Rookies`, is.na(`Rookies`), 0)) %>%
  select(-Rk) %>% 
  mutate(RosterSpace = 17-Teamtotal) %>%
  select(Win, Conf, Div, `Fairly Paid`, Overpaid, Rookies, Underpaid, IR, RosterSpace) %>% 
  arrange(desc(Win))

hhhh <- Save %>%
  left_join(uuuu)


EE <- rpart(Win~ `Fairly Paid`+Overpaid+ Rookies+Underpaid+IR+RosterSpace+PG + SG + SF + PF + C, data = hhhh, method="class", minsplit = 4, minbucket=4) 
summary(EE)
rpart.plot(EE, type = 3, clip.right.labs = FALSE, branch = .3, under = TRUE)
rpart.rules(EE, cover = TRUE)
predict(EE)
round(resid(EE), digits = 0)
```

```{r}

Categorize5Teamsdata <- Save %>% pivot_longer(cols = c('Fairly Paid', 'Overpaid', 'Rookies','Underpaid', 'IR', 'RosterSpace'), names_to = 'Type', values_to = 'Typetotal') %>%
  mutate(WinCat = cut(Win, breaks = c(19, 25,35,43,48,51,90), labels = c("Worst", "WBA", "BA", "AA","BAA", "Best"))) %>%
  group_by(WinCat, Type) %>%
  summarize(Typetotal = mean(Typetotal))
c("1", "2","3","4","5","6")
Categorize5Teamsdata %>%ggplot(aes(y=Typetotal, x=WinCat, fill = Type)) +
  geom_col() +
  labs(caption= 
  "Variable Importance Scores...
  Rookies:30
  Fairly Paid:26
  Underpaid:21
  Overpaid:11
  IR:9
  RosterSpace:2") +
  ylab("Number of Players") +
  theme_classic()

IndividualTeamsdata <- Save %>% pivot_longer(cols = c('Fairly Paid', 'Overpaid', 'Rookies','Underpaid', 'IR', 'RosterSpace'), names_to = 'Type', values_to = 'Typetotal')

IndividualTeamsdata%>%ggplot(aes(y=Typetotal, x=Win, color = Type)) +
  geom_point() +
  geom_smooth(se = FALSE)+
  labs(caption= 
  "Variable Importance Scores...
  Rookies:30
  Fairly Paid:26
  Underpaid:21") +
  ylab("Number of Players") +
  ylim(0, 10) +
  theme_classic() 

write.csv(IndividualTeamsdata, "IndividualTeamsdata.csv")

```


# Trends 6 Categories
```{r}
CatTrend <- Categorize5Teamsdata %>%
  pivot_wider(names_from = `Type`, values_from = `Typetotal`)
CatTrend
```



```{r}
IndividualTeamsdata <- Save %>% pivot_longer(cols = c('Fairly Paid', 'Overpaid', 'Rookies','Underpaid', 'IR', 'RosterSpace'), names_to = 'Type', values_to = 'Typetotal')

IndividualTeamsdata%>%ggplot(aes(y=Typetotal, x=Win, color = Type)) +
  geom_point() +
  geom_smooth(se = FALSE)+
  labs(caption= 
  "Variable Importance Scores...
  Rookies:30
  Fairly Paid:26
  Underpaid:21") +
  ylab("Number of Players 2022-2023") +
  xlab("Team WIns in 2021-2022 Season") +
  ylim(0, 10) +
  theme_classic()

#Individual
#, minsplit = 3, minbucket=4
EE <- rpart(Win~ `Fairly Paid`+Overpaid+ Rookies+Underpaid+IR+`RosterSpace`, data = Save, minsplit = 0, minbucket=4) 
printcp(EE)
plotcp(EE)
summary(EE)
rpart.plot(EE, type = 3, clip.right.labs = FALSE, branch = .5, under = TRUE)
rpart.rules(EE, cover = TRUE)
Save$predic = predict(EE)
Save$residual =round(resid(EE), digits = 1)

ggplot(Save, aes(y = residual, x = Win,color= residual > 0)) + geom_point()
Save
```


```{r}
Current_Team_Rank <- read_csv("C:/Users/ethan/OneDrive/Macalester Sophmore Fall/Introduction to Data Science/Current Team Rank.txt") %>%
  select(-Rk)

```

```{r}
CurrentEverything <- Combo1 %>%
  left_join(Current_Team_Rank)

CurrentSave <- CurrentEverything %>% 
  pivot_wider(names_from = `Type`, values_from = `Typetotal`) %>%
  mutate("IR" = replace(IR, is.na(IR), 0)) %>%
  mutate("Fairly Paid" = replace(`Fairly Paid`, is.na(`Fairly Paid`), 0)) %>%
  mutate("Rookies" = replace(`Rookies`, is.na(`Rookies`), 0)) %>% 
  mutate(RosterSpace = 17-Teamtotal) %>%
  arrange(desc(W))


CurrentIndividualTeamsdata <- CurrentSave %>% pivot_longer(cols = c('Fairly Paid', 'Overpaid', 'Rookies','Underpaid', 'IR', 'RosterSpace'), names_to = 'Type', values_to = 'Typetotal')

CurrentIndividualTeamsdata%>%
  ggplot(aes(y=Typetotal, x=W, color = Type)) +
  geom_point() +
  geom_smooth(se = FALSE)+
  labs(caption= 
  "Variable Importance Scores...
  Rookies:
  Fairly Paid:
  Underpaid:") +
  ylab("Number of Players") +
  xlab("Team Wins in 2022-23 Season as of Dec 12,2022")+
  ylim(0, 10) +
  theme_classic()

#Individual
#, minsplit = 3, minbucket=4
CEE <- rpart(W~ `Fairly Paid`+Overpaid+ Rookies+Underpaid+IR+`RosterSpace`, data = CurrentSave, minsplit = 0, minbucket=4) 
printcp(CEE)
plotcp(CEE)
summary(CEE)
rpart.plot(CEE, type = 3, clip.right.labs = FALSE, branch = .5, under = TRUE)
rpart.rules(CEE, cover = TRUE)
CurrentSave$predic = predict(CEE)
CurrentSave$residual =round(resid(CEE), digits = 1)

ggplot(Save, aes(y = residual, x = Win,color= residual > 0)) + geom_point()
CurrentSave
```
```{r}
FinalDataSet %>%
  mutate("2022-23" = replace(`2022-23`, is.na(`2022-23`), 0)) %>%
  group_by(Team,Pos) %>%
  summarize(teambyposCONTRACT = sum(`2022-23`)) %>%
  group_by(Team) %>%
  mutate(teamCONTRACT = sum(teambyposCONTRACT)) %>%
  ggplot(aes(x=Team, y=teambyposCONTRACT, fill = Pos) ) +
  geom_col() +
  coord_flip()

TeamContractandPos <- FinalDataSet %>%
  mutate("2022-23" = replace(`2022-23`, is.na(`2022-23`), 0)) %>%
  group_by(Team,Pos) %>%
  summarize(teambyposCONTRACT = sum(`2022-23`)) %>% 
  left_join(updateabb_team) %>%
  rename(TeamAbb = Team) %>%
  rename(Team = `Official Team Name`) %>%
  left_join(Current_Team_Rank) %>%
  group_by(Team) %>%
  mutate(teamContract = sum(teambyposCONTRACT)) 
  
  
TeamContractandPos %>% ggplot(aes(x = W, y = teambyposCONTRACT, color = Pos)) + 
  geom_point() +
  geom_smooth(se = FALSE)

TeamContractandPos %>%
  ggplot(aes(x = W, y = teamContract)) + 
  geom_point(alpha=0) +
  geom_smooth() +
  geom_text(aes(label = TeamAbb), size = 3) +
  scale_y_continuous(label=comma)
```

