---
title: "Project Report"
author: "Owen Suelflow, Jonah Klein-Collins, Ethan Caldecott, Harry Cheng"
date: "2022-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE}
# Import the Data
Hustle <- read_csv("/Users/owensuelflow/Documents/COMP_127/NBA_STAT_R_PROJECT/projectdata/hustle.csv")
Boxouts <- read_csv("/Users/owensuelflow/Documents/COMP_127/NBA_STAT_R_PROJECT/projectdata/boxouts.csv")
Drives <- read_csv("/Users/owensuelflow/Documents/COMP_127/NBA_STAT_R_PROJECT/projectdata/drives.csv")
playoff_status <- read_csv("/Users/owensuelflow/Documents/COMP_127/NBA_STAT_R_PROJECT/playoff_status.csv")
nba_advanced_data_4_Sheet1 <- read_csv("/Users/owensuelflow/Documents/COMP_127/NBA_STAT_R_PROJECT/nba advanced data-4 - Sheet1.csv")
Passes <- read_csv("/Users/owensuelflow/Documents/COMP_127/NBA_STAT_R_PROJECT/projectdata/passes.csv")
PnR_Handler <- read_csv("/Users/owensuelflow/Documents/COMP_127/NBA_STAT_R_PROJECT/projectdata/p_r_handler.csv")
PnR_Roller <- read_csv("/Users/owensuelflow/Documents/COMP_127/NBA_STAT_R_PROJECT/projectdata/p_r_roller.csv")
Transitions <- read_csv("/Users/owensuelflow/Documents/COMP_127/NBA_STAT_R_PROJECT/projectdata/transitions.csv")
Salary <- read_csv("/Users/owensuelflow/Downloads/IndividualTeamsdata.csv")
Touches <- read_csv("/Users/owensuelflow/Documents/COMP_127/NBA_STAT_R_PROJECT/projectdata/touches.csv")



Hustle <- Hustle %>%
  mutate(LooseBallsRecovered = `Loose BallsRecovered`)
Hustle <- Hustle %>% 
  mutate(W = Drives$W)
Boxouts <- Boxouts %>% 
  mutate(BoxOuts = `Box Outs`)
Boxouts <- Boxouts %>% 
  mutate(LooseBalls = Hustle$LooseBallsRecovered)
Boxouts <- Boxouts %>% 
  mutate(Playoffs = playoff_status$Playoffs)
nba_advanced_data_4_Sheet1 <- nba_advanced_data_4_Sheet1 %>% 
  arrange(Team)
Boxouts <- Boxouts %>% mutate(Avg_Team_Age = nba_advanced_data_4_Sheet1$Age)
```


## Intro

This report will examine NBA (National Basketball Association) play style data and roster construction to try and answer the following questions: What tendencies and characteristics do playoff teams have? What patterns begin to emerge? The majority of our data was scraped directly off of NBA.com, while a few data sets were downloaded from basketball-reference.com. One thing that we had to be careful of was drawing conclusions based off of reverse causality. This concept occurs when comparing two or more variables that influence one another. For example, comparing point differential among NBA teams, meaning the difference between the average points scored and points given up, and the number of wins each team had. This would show us a clear trend of teams with a higher point differential having more wins. However, by definition a win occurs when a team scores more points than its opponent, so both of out variables in this scenario depend on each other. These are the types of conclusions we strived to avoid. 

The nice thing about play style in the NBA is that different teams have different philosophies regarding how they play. We wanted to look at data that showed a "choice" a team made on how to play. For example, a team can choose to take more drives during a game. On the other hand, a team could decide to avoid driving into the paint and place more of an emphasis on shooting 3 point shots. These types of data do not have reverse causality with wins. 

For roster construction, the line regarding reverse causality is a bit more blurred. Teams may not have a choice between two options when building a team; depending on if a team is contending or rebuilding, they will likely have only one path to follow. However, it is still important to look at roster construction as a tool to follow how teams are doing. We can see if a team that is contending is actually where it should be, or if they should consider pivoting into a rebuild.

## Shiny App

To begin, we performed exploratory data analysis using a shiny app. The app only focused on play style data from eight different data sets, each focusing on one aspect of a game. We were able to choose which data set to look at, and then choose a variable to plot against team wins. This way, we could see if there were any variables with a strong correlation to winning. Here is a link to the shiny app:

https://rstudio-connect.macalester.edu/content/ce590312-5b40-4e4c-84a9-30ff6e7b5a56

## Jonah Viz 1

```{r, echo=FALSE, warning = FALSE, message=FALSE}

```

## Jonah Viz 2

```{r, echo=FALSE, warning = FALSE, message=FALSE}
Transitions %>% 
  left_join(playoff_status, by = c("TEAM" = "Team")) %>% 
  left_join(Touches[,1:3]) %>% 
  ggplot(aes(x = `FREQ%`, y = PPP, color = Playoffs))+
  geom_point(size = 3)+
  scale_color_manual(values = c("brown", "darkgoldenrod2", "chartreuse4"))+
  theme_classic()+
  geom_hline(yintercept = 1.135, alpha = .2)+
  annotate(geom = "text", x = 17.5, y = 1.138, label = "Cutoff Line", size = 2, alpha = .2)+
  labs(y = "Points Per Play", x = "Transition Frequency", title = "Transition Frequency, Efficiency, and Playoff Status")
```

## Owen Viz 1

While there is not a correlation between having lots of transition opportunities and winning, there are other hustle-related statistics that do show a positive correlation to winning. The visualization below shows that playoff/play-in teams tend to win more loose balls and execute more box outs per game than teams who did not make the playoffs. Another interesting trend that we noticed was that a lot of the teams that performed well in these statistics were older. The Miami Heat, for example are one of the oldest teams in the NBA, yet they led the league in box outs. This trend shows up throughout the plot. It could be because older teams are more experienced and generally of higher quality, as we will show later on. These teams know what it takes to win, and oftentimes that means honing in on the smaller details of the games. On the flip side, younger teams generally lose more games, and they aren't disciplined enough to hustle after loose balls or get a body on somebody and get a rebound. Furthermore, these young teams don't have the chemistry and team culture like a team such as the Miami Heat. Overall, it is clear that winning teams prioritize hustle and sound defense, and that experienced teams know these things are crucial to win games in the NBA.

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ggplot(Boxouts, aes(x = BoxOuts, y = LooseBalls, color = Playoffs, size = Avg_Team_Age)) +
  geom_point() +
  geom_text(aes(label = Team), nudge_y = 0.08, size = 2, nudge_x = 0.3) +
  xlim(6.5, 11.4) + 
  scale_color_manual(values = c("brown", "darkgoldenrod2", "chartreuse4")) +
  theme_classic()
```

## Harry Viz 1

```{r, echo=FALSE, warning = FALSE, message=FALSE}
nba_advanced_data5 <- nba_advanced_data_4_Sheet1 %>% 
  left_join(playoff_status) 

nba_advanced_data5 %>% 
select(Team,Age,W,Playoffs) %>% 
  ggplot(aes(x=Age,y=W))+geom_text(aes(label=Team, color = Playoffs), size=2)+geom_smooth(se=FALSE,method="gam")+xlim(21,32)+
  scale_color_manual(values = c("brown", "darkgoldenrod2" , "chartreuse4"))+
  theme_classic()
```

## Harry Viz 2

```{r, echo=FALSE, warning = FALSE, message=FALSE}
nba_advanced_data5 %>% 
select(Team,Age,ORtg,Playoffs) %>% 
  ggplot(aes(x=Age,y=ORtg))+geom_text(aes(label=Team, color = Playoffs), size=2)+geom_smooth(se=FALSE,method = "gam")+xlim(21,32)+
  scale_color_manual(values = c("brown", "darkgoldenrod2" , "chartreuse4"))+theme_classic()
```

## Ethan Viz 1

```{r, echo=FALSE, warning = FALSE, message=FALSE}

```

## Ethan Viz 2

```{r, echo=FALSE, warning = FALSE, message=FALSE}

```

## Ethan Viz 3

```{r, echo=FALSE, warning = FALSE, message=FALSE}

```

## Conclusion